ggalluvial: Layered Grammar for Alluvial Plots

### 2020-03-22 paper.rmd
clarify labels in joss paper figure

```{r figure, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=8, fig.cap="\\label{fig:example}Alluvial plot of changes in curricula in a student cohort"}
library(ggalluvial)
data(majors)
majors$curriculum <- as.factor(majors$curriculum)
ggplot(majors,
       aes(x = semester, stratum = curriculum, alluvium = student,
           fill = curriculum)) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  scale_y_continuous(breaks = seq(-2, 10, 2)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback",
            color = "darkgray", negate.strata = NA) +
  geom_stratum(negate.strata = NA) +
  theme(legend.position = "right") +
  labs(x = NULL)
```

>>>>

```{r figure, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=8, fig.cap="\\label{fig:example}Alluvial plot of changes in curricula by a cohort of art students"}
library(ggalluvial)
data(majors)
majors$curriculum <- as.factor(majors$curriculum)
levels(majors$semester) <- gsub("CURR", "Term ", levels(majors$semester))
arr <- arrow(angle = 15, type = "closed", length = unit(.03, "npc"))
ggplot(majors,
       aes(x = semester, stratum = curriculum, alluvium = student,
           fill = curriculum)) +
  scale_fill_brewer(type = "qual", palette = "Set2", name = "Curriculum") +
  scale_y_continuous(breaks = seq(-2, 10, 2)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback",
            color = "darkgray", negate.strata = NA) +
  geom_stratum(negate.strata = NA) +
  theme(legend.position = "right") +
  labs(x = NULL, y = "Number of students") +
  # annotate axes
  annotate("segment", x = 2.33, y = -1.33, xend = 2, yend = -.33, arrow = arr) +
  annotate("segment", x = 3, y = -1.33, xend = 3, yend = -.33, arrow = arr) +
  annotate("segment", x = 3.67, y = -1.33, xend = 4, yend = -.33, arrow = arr) +
  annotate("text", x = 3, y = -1.67, label = "axes / dimensions", size = 4) +
  # annotate strata
  annotate("segment", x = 1.5, xend = 1, y = 3.5, yend = 1.5, arrow = arr) +
  annotate("segment", x = 1.5, xend = 1, y = 4, yend = 3.5, arrow = arr) +
  annotate("segment", x = 1.5, xend = 1, y = 4.5, yend = 6, arrow = arr) +
  annotate("text", x = 1.67, y = 4, label = "strata", angle = 90, size = 4) +
  # annotate alluvia
  annotate("segment", x = 4.33, y = 6.5, xend = 1, yend = 8.5, arrow = arr) +
  annotate("segment", x = 5, y = 6.5, xend = 4.5, yend = 9, arrow = arr) +
  annotate("segment", x = 5.67, y = 6.5, xend = 7.5, yend = 7, arrow = arr) +
  annotate("text", x = 5, y = 6.17, label = "one alluvium", size = 4) +
  # annotate flows
  annotate("segment", x = 5.33, y = 2.5, xend = 4.25, yend = 0.5, arrow = arr) +
  annotate("segment", x = 5.33, y = 3, xend = 4.25, yend = 2.5, arrow = arr) +
  annotate("segment", x = 5.33, y = 3.5, xend = 3.75, yend = 4.5, arrow = arr) +
  annotate("text", x = 5.5, y = 3, label = "flows", angle = 90, size = 4) +
  # annotate lodes
  annotate("segment", x = 7.5, y = 3, xend = 8, yend = 1.5, arrow = arr) +
  annotate("segment", x = 7.5, y = 3.5, xend = 8, yend = 3.5, arrow = arr) +
  annotate("segment", x = 7.5, y = 4, xend = 8, yend = 5.5, arrow = arr) +
  annotate("text", x = 7.33, y = 3.5, label = "lodes", angle = -90, size = 4)
```
